---
title: "Prince George's County Crime Data"
---


```{r}
pg_crimes <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/Crime_Incidents_July_2023_to_Present_20250316.csv") |> clean_names()

pg_crimes <- pg_crimes |> 
  mutate(clean_date = mdy_hms(date), month = month(clean_date), week_day = wday(clean_date))

pg_crimes |> 
  filter(str_detect(clearance_code_inc_type, "AUTO  STOLEN")) |> 
  group_by(month) |> 
  summarize(total = n())
```

### Stolen cars by month

Null Hypothesis: September's decline was not statistically significant and a result of random chance.
Alternative Hypothesis: September's decline was statistically significant.

### Set up our t.test

```{r}

# Load the data for all months from July through February, skipping March because it's not finished
car_theft_totals <- pg_crimes |> 
  filter(str_detect(clearance_code_inc_type, "AUTO  STOLEN")) |> 
  filter(month %in% c(7,8,9,10,11,12,1,2)) |> 
  group_by(month) |> 
  summarize(count = n())
car_thefts <- car_theft_totals$count

# set the confidence interval
alpha <- 0.05

# Calculate the mean and standard deviation of all the months
all_months_mean <- mean(car_thefts)
all_months_sd <- sd(car_thefts)

# Perform the t-test
result <- t.test(car_thefts, mu = all_months_mean, alternative = "less")

# Print the results
print(result)

# Calculate p-value
p_value <- result$p.value

# Make a decision
if (p_value < alpha) {
  decision <- "reject the null hypothesis"
} else {
  decision <- "fail to reject the null hypothesis"
}

# Interpretation
interpretation <- paste("Based on the p-value of", round(p_value, 4),
                        ", at a significance level of", alpha,
                        ", we", decision)

interpretation
```

#### What about murders?

Null Hypothesis: There has been no significant change in murders since last summer.
Alternative Hypothesis: There has been a significant change in murders since last summer.

### Set up our data

```{r}
# Given data
homicides_month <- pg_crimes |> 
  filter(month %in% c(7,8,9,10,11,12,1,2)) |>   # March isn't finished yet
  filter(clearance_code_inc_type == 'HOMICIDE') |> 
  group_by(month) |> 
  summarize(total = n())

# Set significance level
alpha <- 0.05

# Load the data from July through February
homicides <- homicides_month$total

# Calculate the mean and standard deviation of the data
homicides_mean <- mean(homicides)
homicides_sd <- sd(homicides)

# Perform the t-test
results <- t.test(homicides, mu = homicides_mean)

# Print the results
print(results)

# Extract p-value
p_value <- results$p.value

# Make decision based on p-value
if (p_value < alpha) {
  decision <- "reject the null hypothesis"
} else {
  decision <- "fail to reject the null hypothesis"
}

decision

```
This result indicates that the variation in homicides across the 9 months, from July to February, is not statistically significant. The monthly homicide counts are not different from the overall mean of 5.77 in a statistically meaningful way. The t-test shows that the fluctuations in homicides during this time period are likely just due to random chance and do not represent a significant change or pattern in the data. The ranges for the 95 percent confidence interval show this; 7 is within that range.
